---
title: "Analyzing Macroeconomic Trends with Aurora"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing Macroeconomic Trends with Aurora}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `aurora` package allows you to easily fetch and align disparate data sources—weather, economics, and events—into a single daily timeline.

## Usage

```{r setup}
library(aurora)
```

### Fetching Data

You can fetch a comprehensive dataset for a specific location and date range. Here we look at Ottawa, Canada for January 2024.

```{r fetch, eval = FALSE}
# Fetches data from Open-Meteo, FRED, and GDELT
data <- get_macroeconomic_data(
  start_date = "2024-01-01",
  end_date = "2024-01-31",
  latitude = 45.4215,
  longitude = -75.6972
)

head(data)
```

### Visualizing Trends

Once you have the data, you can easily plot relationships using `ggplot2`.

```{r plot, eval = FALSE}
library(ggplot2)

ggplot(data, aes(x = DATE)) +
  geom_line(aes(y = avg_temp, color = "Temperature")) +
  geom_line(aes(y = political_unrest_score, color = "Political Unrest")) +
  labs(title = "Temperature vs Political Unrest", y = "Value")
```
